---
title: "Use a storage"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Use a storage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)

Sys.setlocale("LC_ALL", "English")

```

```{r eval=FALSE, message=FALSE, warning = FALSE}

library(barrks)

```

A typical application of `barrks` is to repeatedly calculate a phenology with data
on a daily basis. Depending on the model and the number of raster cells, these
calculations can be quite time-consuming. As most of the calculations are redundant
to earlier versions, the use of a storage can greatly reduce the computation time.
A storage can be used by passing a directory as `.storage` parameter to `phenology()`.
A phenology can be recovered from a storage by
calling `phenology` with `.storage` parameter and without model inputs.

Note that the use of a storage is neither available for station data nor for the `onset`
and the `development` submodels of the BSO model. As saving and loading from
the storage takes time as well, the calculation of a phenology might occasionally be slower,
especially for small rasters.

```{r, eval=FALSE}

# call phenology with a storage path
pheno <- phenology('phenips-clim',
                   barrks_data(),
                   .storage = 'path/to/storage')

# recover phenology from a storage path
pheno_recovered <- phenology('phenips-clim', .storage = 'path/to/storage')

```

The functions `save_phenology()` and `load_phenology()` provide a faster way for
saving and loading phenology objects, but they do not allow successive calculations.


